---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# arnetworks

The goal of arnetworks is to simulate, estimate and predict from autoregressive (AR) network models with edge dependence, following the framework specified in Fang et al. (2024+).

## Installation

You can install the development version of arnetworks from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
# devtools::install_github("peterwmacd/arnetworks")
```

## Example 1: Transitivity Model

The package provides a detailed implementation for fitting a particular AR network model with transitivity effects (see Fang et al. (2024+), Section X.X). This is a basic example which shows how to simulate, estimate and predict with this model.

```{r example1}
library(arnetworks)
# transitivity model


```

## Example 2: Persistence Model

The package also allows users to specify their own AR network models with both local and global parameters. This is a basic example which shows how to simulate, estimate and predict with the persistence model (see Fang et al. (2024+), Section Y.Y).

```{r example2}
library(arnetworks)
# Persistence Model

# Set model parameters
p = 30; n = 20
xi = runif(p, 0.5, 0.9)
eta = runif(p, 0.5, 0.9)
a = 0.5
b = 0.5
# Simulation with simulatePersistence
data2 = simulatePersistence(p, n, xi, eta, a, b)

# Specify model components for fitting with EstNet

# Specify data and sufficient statistics
X = data2$X
statsAlpha =  1-X[,,4:n-2]+ ( 1-X[,,4:n-2])*( 1-X[,,4:n-3])
statsBeta =  X[,,4:n-2]+ ( X[,,4:n-2])*( X[,,4:n-3])
X = X[,, 3: n]
# NOTE: estimation implicitly conditions on the first two network snapshots

# Define edge formation and dissolution functions
fij = function(global, stats) {
  return (exp(-1 -global*stats))
}
gij <- function(global, stats) {
  return (exp(-1 -global*stats))
}
# Model fitting with estNet
fit2 <- estNet(X, fij, gij, 
                 statsAlpha, statsBeta, 
                 globInitAlpha=1, globInitBeta=1,
                 shrGPrm = 0, maxIter = 100)

# global parameter for edge formation
print(fit2$gAlphaVal)
# global parameter for edge dissolution
print(fit2$gBetaVal)
# local parameters for edge formation (quartiles)
print(quantile(fit2$xi))
# local parameters for edge formation (quartiles)
print(quantile(fit2$eta))

# Predict the next network snapshot

# Specify snapshot to predict from
Xnew = data2$X[,,n]
# Specify the sufficient statistics for prediction (as arrays)
statsAlphaNew =  1-data2$X[,,n-1,drop=FALSE]+ ( 1-data2$X[,,n-1,drop=FALSE])*( 1-data2$X[,,n-2,drop=FALSE])
statsBetaNew =  data2$X[,,n-1,drop=FALSE]+ ( data2$X[,,n-1,drop=FALSE])*( data2$X[,,n-2,drop=FALSE])
# Prediction with predictNet
pred2 = predictNet(fit2,Xnew,statsAlphaNew,statsBetaNew,fij,gij)

```
